### 2021-05-21 concurrency/docs/theory/theory.tex
typo
another typo
make something more clear
start drafting the formalized rwlock example

Now we define some functions on this state, defined recursively.
First, for any refinement location $\delta \refines_{\rel} \gamma$,
where $\rel : M \times N \to \bool$,
we let $Project_\delta : M \to N$ be a ``functionalization'' of $\rel$. In particular,
\[
  \forall m .~ domain(\rel, m) \implies \rel(m, Project_\delta(m)).
\]
where $domain(\rel, m) \eqdef \exists n .~ \rel(m, n)$.

@@

Thus we will have an invariant, only applying to the active lifetime $active$,
\begin{align}
  domain(rel, Total_{active}(\delta \refines_{\rel} \gamma)) \label{invTotal}
\end{align}

@@

Again, we define a ``projection'' function from a view on a refinement TPCM to a base TPCM.
\[
  ProjectView_\delta(v) \eqdef \lambda a .~
      \exists x, y .~ v(x) \land rel(x, y) \land y \le a.
\]
For a given $m$, we define $ViewOver(m)$ to be the view of all elements that are at least $m$:
\[
  ViewOver(m) \eqdef \lambda n .~ n \ge m
\]
And we can add two views in the natural way:
\[
  v \otimes w \eqdef \lambda m .~ \exists x, y .~ v(x) \land w(y) \land x \cdot y = m
\]
Finally, we're ready to define the $View_\kappa$ function for a given location, again
recursively, like the $Total_\kappa$ function:
\[
  View_\kappa(\gamma) \eqdef 
      \left(
        \bigotimes_{\kappa' \ltGe \kappa} ViewOver(reserved(\kappa,\gamma))
      \right) \otimes
      \left(
        \bigotimes_{\delta} ProjectView_\delta\left(View_\kappa(\delta \refines_{\rel} \gamma)\right)
      \right)
\]
Now the $View_\kappa$ denotes the ``possible'' states that the total unlive state can be in
for a given $\gamma$, without actually depending on any live state even in the refining locations. The $View_\kappa$ is in some sense used to justify the borrows at $\kappa$, so we will have an invariant,
\begin{align}
  & \forall \kappa, \gamma, m .~
  (\kappa, \gamma, m) \in borrows \implies \forall y .~ View_\kappa(\gamma)(m)(y) \implies m \le y
  \label{invView}
\end{align}
    %\left(
    %    \Bigcircle_{\kappa' \ltGe \kappa} reserved(\kappa,\gamma)
    %  \right) \cdot
    %  \left(
    %    \Bigcircle_{\delta} Project_\delta\left(Total_\kappa(\delta \refines_{\rel} \gamma)\right)
    %  \right)
   %\right)
%\end{align}
(We don't need a condition that $\kappa \ltGe active$, although we won't be able to make use
of the invariant otherwise.)

@@

Second, we have to show why it is allowed to use borrowed state to justify the transitions.
Luckily, from (\ref{invTotal}) we can derive this law,
\[
  \kappa \ltGe active \implies View_\kappa(\gamma)(y)\left(
      \left(
        \Bigcircle_{\kappa' \ltGe active} reserved(\kappa,\gamma)
      \right) \cdot
      \left(
        \Bigcircle_{\delta} Project_\delta\left(Total_{active}(\delta \refines_{\rel} \gamma)\right)
      \right)
    \right).
    \label{invDerived}
\]
This large expression inside the $View_\kappa(\gamma)$ is just the part of
$Total_{active}(\gamma)$ except the part \emph{which is live at $\gamma$}.
Thus, combined with \ref{invView}, we can see that any borrow will be covered by the reserved 
state at $\gamma$ plus all state in the refined locations.

@@

>>>>

Now we define some functions on this state, defined recursively.
First, for any refinement location $\delta \refines_{\rel} \gamma$,
where $\rel : M \times N \to \bool$,
we let $Project_\delta : M \to N$ be a ``functionalization'' of $\rel$. In particular,
\[
  \forall m .~ m \in domain(\rel) \implies \rel(m, Project_\delta(m)).
\]
where $domain(\rel) \eqdef \{m : \exists n .~ \rel(m, n)\}$.

@@

Thus we will have an invariant, only applying to the active lifetime $active$,
\begin{align}
  Total_{active}(\delta \refines_{\rel} \gamma) \in domain(\rel) \label{invTotal}
\end{align}

@@

Again, we define a ``projection'' function from a view on a refinement TPCM to a base TPCM.
\[
  ProjectView_\delta(v) \eqdef \lambda a .~
      \exists x, y .~ v(x) \land y = Project_\delta(x) \land (x \in domain(rel)) \land y \le a.
\]
For a given $m$, we define $ViewOver(m)$ to be the view of all elements that are at least $m$:
\[
  ViewOver(m) \eqdef \lambda n .~ n \ge m
\]
And we can add two views in the natural way:
\[
  v \otimes w \eqdef \lambda m .~ \exists x, y .~ v(x) \land w(y) \land x \cdot y = m
\]
Finally, we're ready to define the $View_\kappa$ function for a given location, again
recursively, like the $Total_\kappa$ function:
\[
  View_\kappa(\gamma) \eqdef 
      \left(
        \bigotimes_{\kappa' \ltGe \kappa} ViewOver(reserved(\kappa,\gamma))
      \right) \otimes
      \left(
        \bigotimes_{\delta} ProjectView_\delta\left(View_\kappa(\delta \refines_{\rel} \gamma)\right)
      \right)
\]
Now the $View_\kappa$ denotes the ``possible'' states that the total unlive state can be in
for a given $\gamma$, without actually depending on any live state even in the refining locations. The $View_\kappa$ is in some sense used to justify the borrows at $\kappa$, so we will have an invariant,
\begin{align}
  & \forall \kappa, \gamma, m .~
  (\kappa, \gamma, m) \in borrows \implies \forall y .~ View_\kappa(\gamma)(y) \implies m \le y
  \label{invView}
\end{align}
    %\left(
    %    \Bigcircle_{\kappa' \ltGe \kappa} reserved(\kappa,\gamma)
    %  \right) \cdot
    %  \left(
    %    \Bigcircle_{\delta} Project_\delta\left(Total_\kappa(\delta \refines_{\rel} \gamma)\right)
    %  \right)
   %\right)
%\end{align}
(We don't need a condition that $\kappa \ltGe active$, although we won't be able to make use
of the invariant otherwise.)

@@

Second, we have to show why it is allowed to use borrowed state to justify the transitions.
Luckily, from (\ref{invTotal}) we can derive this law,
\[
  \kappa \ltGe active \implies View_\kappa(\gamma)\left(
      \left(
        \Bigcircle_{\kappa' \ltGe active} reserved(\kappa,\gamma)
      \right) \cdot
      \left(
        \Bigcircle_{\delta} Project_\delta\left(Total_{active}(\delta \refines_{\rel} \gamma)\right)
      \right)
    \right).
    \label{invDerived}
\]
This large expression inside the $View_\kappa(\gamma)$ is just the part of
$Total_{active}(\gamma)$ except the part \emph{which is live at $\gamma$}.
Thus, combined with \ref{invView}, we can see that any borrow will be covered by the reserved 
state at $\gamma$ plus all state in the refined locations.

@@

\textbf{Worked RWLock example}

\newcommand{\Central}{\textsf{Central}}
\newcommand{\ExcPending}{\textsf{ExcPending}}
\newcommand{\SharedPending}{\textsf{SharedPending}}
\newcommand{\ExcGuard}{\textsf{ExcGuard}}
\newcommand{\SharedGuard}{\textsf{SharedGuard}}

Given a TPCM $M$, we can construct a TPCM $RwLock_M$ with the following laws. (Here, we have $exc : \bool$, $rc \in \mathbb{N}$, and $x \in M$.

These laws are for exclusive locks:
\begin{align*}
\dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma} &\viewshift \dboxed{\Central\left(\true, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma}\\
%\Central\left(exc, 0, x\right) \cdot \ExcPending &\mapsto \Central\left(exc, 0, \bot\right) \cdot \ExcGuard
\dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma}
    &\viewshift
    \dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma\\
\dboxed{\Central\left(exc, rc, y\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma
    &\viewshift
        \dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma}\\
\end{align*}
The first two are for acquisition, while the third is for release. Meanwhile,
the first one is derived from Ghost-ViewShift, while the last two are derived from
Ghost-Refinement-Exchange.

These laws are for shared locks:
\begin{align*}
\dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} &\viewshift
  \dboxed{\Central\left(exc, rc+1, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}\\
\dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}
    &\viewshift
      \dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}\\
\dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}
    &\viewshift
        \dboxed{\Central\left(exc, rc - 1, x\right)}^{\alpha \refines \gamma}
\end{align*}
All of these are derived from Ghost-ViewShift.

We don't use Ghost-Refinement-Exchange, because in the `shared' case, our mechanism for getting back $M$-state is to use the Borrow-Back rule. In particular, we use it to derive this:
\begin{align*}
\borrow{\kappa}{\dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}}
  &\viewshift
    \borrow{\kappa}{\dboxed{x}^{\gamma}}
\end{align*}
This logic is all we need to verify a reader-writer lock.
